Conversion of FFCS ICE fleet to EV: a data-driven approach

### 2020-11-16 Chapter10/10_1-introduction.tex
chap_9_parsed

%****************************************
%What we did in this work
%***************************************
In this work we analyze the feasibility and scalability of FFCS with electric vehicles.
Our goal is to find an economically sustainable solution that brings benefits to both citizens (i.e., high availability) and operators (i.e., high profit), with the economic sustainability being a crucial aspect. For instance in Italy, the main FFCS operator had revenues around 48 million euros in 2016, but still losing around 27 millions euros, with each car burning 4700\,€ on average\cite{Quattroruote}.
Despite that, car sharing is estimated to increase from 20\% to 40\% from 2019 to 2021 \cite{soleCarSharing}. %\lv{'Il Sole 24 Ore' e la citazione, ma ci serve o url o il numero esatto, e la pagina. Chi l'ha inserita?}.

@@

As a case study, we focus on the city of Turin in Italy. 
We leverage hundred of thousands of real FFCS trips~\cite{10_ciociola2017umap} to extract the geo-temporal mobility demand. We use Kernel Density Estimation (KDE) to catch the demand spatial variability, and modulated Poisson models for the temporal demand~\cite{ciociola2020}. 
We use it to feed a trace-driven flexible simulator that allows us to study how the design choices and system parameters impact on performance. We first consider an electric-car sharing system that has the same number of cars and faces same demand of the current one in Turin. We observe the impact of different charging infrastructure design, i.e., the number of poles and how to spread these are over the city area.

Next, we consider when the intensity of the mobility demand grows. How would the charging infrastructure need to grow correspondingly? And what is the impact of the fleet size? 
Summarizing our contributions, this paper proposes an answer to these questions making use of the demand model to project future or different scenarios and the cost-revenue model to evaluate the profitability of each configuration.
We focus on performance indicators like the fraction of demand the system can satisfy and the total working hours it has to spend for the battery charging operations. 
Then, we project these into economical figures, observing how the design options impact on profitability.

%****************************************
%Results and contributions
%***************************************
Our results show that the charging station placement is fundamental if poles are placed in areas with high demand, as cars get located where customers need them. This allows the system to naturally intercept the customers demand, thus to maximize the satisfied demand, and revenues.
Considering system scalability, as expected the charging infrastructure must grow proportionally to the mobility demand. Interestingly instead, the number of vehicles can grow much slower, showing economy of scale savings which make the system likely profitable if well designed.

%****************************************
%Organization of the paper
%***************************************
The paper is organized as follows: In Section~\ref{sec:10_2_related_work} we discuss our work in light of past literature for FFCSs and their economical aspects. After reporting the details about the dataset and demand model in  Section~\ref{sec:10_3_dataset}, and our simulator in Section~\ref{sec:10_4_ModelSimulation}, we present results in Section~\ref{sec:10_5_results} before drawing conclusions in Section~\ref{sec:10_6_conclusions}.

>>>>

%****************************************
%What we did in this work
%***************************************
In this work I analyze the feasibility and scalability of FFCS with electric vehicles. The goal is to find an economically sustainable solution that brings benefits to both citizens (i.e., high availability) and operators (i.e., high profit), with the economic sustainability being a crucial aspect. For instance in Italy, the main FFCS operator had revenues around 48 million euros in 2016, but still losing around 27 millions euros, with each car burning 4700\,€ on average\cite{Quattroruote}.
Despite that, car sharing is estimated to increase from 20\% to 40\% from 2019 to 2021 \cite{soleCarSharing}. %\lv{'Il Sole 24 Ore' e la citazione, ma ci serve o url o il numero esatto, e la pagina. Chi l'ha inserita?}.

@@

As a case study, I focus on the city of Turin in Italy. 
I leverage hundred of thousands of real FFCS trips~\cite{10_ciociola2017umap} to extract the geo-temporal mobility demand. I use Kernel Density Estimation (KDE) to catch the demand spatial variability, and modulated Poisson models for the temporal demand~\cite{ciociola2020}. 
I use it to feed a trace-driven flexible simulator that allows us to study how the design choices and system parameters impact on performance. I first consider an electric-car sharing system that has the same number of cars and faces same demand of the current one in Turin. I observe the impact of different charging infrastructure design, i.e., the number of poles and how to spread these are over the city area.

Next, I consider when the intensity of the mobility demand grows. How would the charging infrastructure need to grow correspondingly? And what is the impact of the fleet size? 
Summarizing the contributions, this paper proposes an answer to these questions making use of the demand model to project future or different scenarios and the cost-revenue model to evaluate the profitability of each configuration.
I focus on performance indicators like the fraction of demand the system can satisfy and the total working hours it has to spend for the battery charging operations. 
Then, I project these into economical figures, observing how the design options impact on profitability.

%****************************************
%Results and contributions
%***************************************
The results show that the charging station placement is fundamental if poles are placed in areas with high demand, as cars get located where customers need them. This allows the system to naturally intercept the customers demand, thus to maximize the satisfied demand, and revenues.
Considering system scalability, as expected the charging infrastructure must grow proportionally to the mobility demand. Interestingly instead, the number of vehicles can grow much slower, showing economy of scale savings which make the system likely profitable if well designed.

%****************************************
%Organization of the paper
%***************************************
The paper is organized as follows: In section~\ref{sec:10_2_related_work} I discuss this work in light of past literature for FFCSs and their economical aspects. After reporting the details about the dataset and demand model in  section~\ref{sec:10_3_dataset}, and the simulator in section~\ref{sec:10_4_ModelSimulation}, I present results in section~\ref{sec:10_5_results} before drawing conclusions in section~\ref{sec:10_6_conclusions}.

### 2020-11-16 Chapter10/10_2-related.tex
chap_9_parsed

Recently, the availability and abundance of data helped shaping FFCS users' behaviour.  Considering this, some works like~\cite{9_habibi2017comparison} and~\cite{10_ciociola2017umap} scraped data from real ICE FFCS, characterized their services and proposed model generalizations. %In particular both showed how median travel distance and time strongly depend on city conformation. 
Big Data approaches helped researchers to improve the simulation fidelity. In particular, authors of~\cite{0_tachet2017scaling} used data to predict the shareability of an urban ride, finding that this a property city-invariant.  %Surprisingly it depends only trips-related characteristics like trips per hour or average traffic speed and not the city shape. The formula is tested in several cities like San Francisco, Vienna, New York and Singapore, producing accurate results. 
On the same optic, the authors of~\cite{22_li2016design} use data from several American cities to optimize the position of the charging stations of a one-way car sharing, finding that  the optimal  results place the stations in high-demand areas.  
We confirmed this result in \cite{7_cocca2019free} and \cite{8_cocca2019free} where we used a simulation based approach to measure the impact of different design options of an EVs FFCS system. %The studies confirmed that best results come from placing charging station high-demand areas. 
Here use big data to derive realistic demand models to feed accurate simulations. We move one step forward - showing that this is beneficial also as a proxy of relocation, i.e., cars get naturally relocated to high demand zones.

%****************************************
%Economic sustainability
%***************************************
%\subsection{Economic sustainability}
The economic sustainability is another key aspect of car sharing - especially with EVs. Authors of~\cite{20_hua2019joint} studied the economic sustainability of one way electric car sharing systems finding out that charging station should have an amortization period of at least 5 years to produce profits.
The authors of~\cite{19_lemme2019optimization} compared how FFCS with EVs and ICE can produce profits, observing the best compromises with ICE fuelled with cheap and cleaner fuel like ethanol. 
Another study~\cite{21_vasconcelos2017financial} concerning the city of Lisbon found out that switching to EVs would cost more than ICE and would lead to a negative profit of about one million per year. This is largely due to the higher cost of electric cars and of the charging infrastructure.
In our work we explore which are the most efficient and economically sustainable combinations of fleet size, charging infrastructure design, also in light of demand growth.

%****************************************
%Difference of our work
%***************************************
To the best of our knowledge, our work is among the first to study the scalability of a FFCS system with electric vehicles, exploring key parameters like number of poles, fleet size and increase in demand can affects economic and performance of the system.

>>>>

Recently, the availability and abundance of data helped shaping FFCS users' behaviour.  Considering this, some works like~\cite{9_habibi2017comparison} and~\cite{10_ciociola2017umap} scraped data from real ICE FFCS, characterized their services and proposed model generalizations. %In particular both showed how median travel distance and time strongly depend on city conformation. 
Big Data approaches helped researchers to improve the simulation fidelity. In particular, authors of~\cite{0_tachet2017scaling} used data to predict the shareability of an urban ride, finding that this a property city-invariant.  %Surprisingly it depends only trips-related characteristics like trips per hour or average traffic speed and not the city shape. The formula is tested in several cities like San Francisco, Vienna, New York and Singapore, producing accurate results. 
On the same optic, the authors of~\cite{22_li2016design} use data from several American cities to optimize the position of the charging stations of a one-way car sharing, finding that  the optimal  results place the stations in high-demand areas.  
This results in this chapters confirms the results in \cite{7_cocca2019free} and \cite{8_cocca2019free} where I used a simulation based approach to measure the impact of different design options of an EVs FFCS system. %The studies confirmed that best results come from placing charging station high-demand areas. 
Here, the use of big data to derive realistic demand models to feed accurate simulations. I move one step forward - showing that this is beneficial also as a proxy of relocation, i.e., cars get naturally relocated to high demand zones.

%****************************************
%Economic sustainability
%***************************************
%\subsection{Economic sustainability}
The economic sustainability is another key aspect of car sharing - especially with EVs. Authors of~\cite{20_hua2019joint} studied the economic sustainability of one way electric car sharing systems finding out that charging station should have an amortization period of at least 5 years to produce profits.
The authors of~\cite{19_lemme2019optimization} compared how FFCS with EVs and ICE can produce profits, observing the best compromises with ICE fuelled with cheap and cleaner fuel like ethanol. 
Another study~\cite{21_vasconcelos2017financial} concerning the city of Lisbon found out that switching to EVs would cost more than ICE and would lead to a negative profit of about one million per year. This is largely due to the higher cost of electric cars and of the charging infrastructure.
This chapter explores which are the most efficient and economically sustainable combinations of fleet size, charging infrastructure design, also in light of demand growth.

%****************************************
%Difference of our work
%***************************************
To the best of I can tell, this work is among the first to study the scalability of a FFCS system with electric vehicles, exploring key parameters like number of poles, fleet size and increase in demand can affects economic and performance of the system.

### 2020-11-16 Chapter10/10_3-dataset.tex
chap_9_parsed

%****************************************
% Motivation and outline
%****************************************
In this work we leverage the data collected in our previous work~\cite{10_ciociola2017umap} capturing real trips performed by car2go users. These data let us model the users' mobility demand in time and in space. From this, we derive a demand model that generalize the users' demand observed in the real data. We use it to generate realistic traces describing possible user trips and feed them to our event-based simulator to derive performance figures.

@@

\subsection{Dataset}

Our data consist in actual rentals performed by car2go users in Turin. Each observed rental has precise geo-spatial coordinates for trip origin and destination, and accurate timestamps. In Figure~\ref{fig:10_3_rentals_per_day} we summarize the number of daily rentals from June 2017 to January 2018 in Turin, our reference dataset\footnote{For some periods we did not record data due to server failures.}.  Despite the fact that rentals are non-stationary, especially during periods like August and Christmas holidays, the service usage follows a hourly and weekly pattern, not showing any particular growth. From this data, we select three months - from October 1st to December 31st.
Table~\ref{tab:10_3_rental_data} outlines the main characteristics of this data. 400 cars were available, traveling on average less than 4\,km in each trip, for an average rental time of 21 minutes.

To catch customers' habits, in Figure~\ref{fig:10_3_rentals_per_hour} we detail the average number of rentals per hour, separately per working-days (Monday to Friday) and per weekends. The weekdays hourly profile reflects the commuting pattern, with two clear peaks in the morning and evening rush hours. Conversely, during the weekends the number of night rentals is higher, likely due to more nightlife, while the morning peak is drastically smoothed.
Not shown here for lack of space, our data allow us to observe the spatial diversity, with different origin and destination areas over the city. This shows how fundamental is to use actual data to build realistic scenarios for accurate system analysis.

\begin{table}[t]
\begin{center}
\setlength\tabcolsep{5pt} % default value: 6pt reduce cell padding
\caption{Main characteristics of our dataset, recorded in Turin from October to December 2017. Rental time and rental distance report both median (Med) and average (Avg) values.}
\begin{tabular}{|l|l|l|c|c|c|l|}
\hline
\!\multirow{3}{*}{Rentals}\!& \multirow{2}{*}{Fleet} &  \multicolumn{2}{|c|}{Rental}  &  \multicolumn{2}{|c|}{Rental}  &  \!\multirow{3}{*}{Zones}\!\\ 
& \multirow{2}{*}{Size}  & \multicolumn{2}{|c|}{Time~[min]}  & \multicolumn{2}{|c|}{Dist.~[km]} &    \\ \cline{3-6} %&  \multirow{2}{*}{Size[GB]}
 &  &\!Avg\!&\!Med\!&\!Avg\!&\!Med\!& \\ \hline
\hline
 180k & 400 & 21 & 20 & 3.96 & 3.36 & 279 \\ \hline
\end{tabular}
\label{tab:10_3_rental_data}
\end{center}
\end{table}

@@

%****************************************
% Demand model
%****************************************
\subsection{Demand Model}

While we could directly use the original trace to observe system performance, we need a model to observe what-if scenarios, e.g., to observe the impact of a growth in the demand. For this, we use the data at our disposal to create a generalized demand model. We follow the approach we presented in~\cite{ciociola2020}. In a nutshell we model the demand in time by using modulated Poisson processes - a common accepted model for independent service requests of a very large population~\cite{poisson_inh}. To capture the spatial heterogeneity, we generalize the traces using Kernel Density Estimation (KDE)~\cite{kde_spatial}. KDE gives us the possibility to smooth the real data over a multi-dimensional space while maintaining the origin/destination correlation.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%Model of departure times
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
In more details, for the request arrival time process we assume that the inter-arrival time of trips follows an exponential distribution with rate depending on the type (weekend or working day) and hour of the day. We consider 24 time bins of 1\,h each - 48 periods in total. In each time/day bin, the Poisson arrival rate matches the average rate of requests in that time bin in the original dataset as shown in Figure~\ref{fig:rentals_per_hour}. This temporal model allows to scale the overall demand by introducing a global scaling factor $\lambda$ as a multiplier of the request rate of each time bin.

To model the spatial diversity of the demand, we divide the city in a set $Z$ of contiguous 500\,m x 500\,m zones, obtaining in total 279 zones. Each couple of spatial coordinates in the city area  ($x,y$) maps to one and only one zone.  Since each trip $i$ departs from a certain zone (origin $O_i$, described by two coordinates) and arrives to another zone (destination $D_j$, described by two coordinates), it is therefore characterized by two couples of coordinates, that can be represented as 4 scalars. 
For each time bin, we derive an origin and destination matrix counting how many trips were originated from a given zone $O$ and destined to a given zone $D$. Thus, in order to model the OD matrix in each temporal slot, we fit a 4-dimensional KDE based on the aforesaid coordinates.
For each of these matrices, we compute a KDE model, using Gaussian kernels, with bandwidth equal to 1. % , i.e., concentrating 66\% of the body of the distribution  in the same 500\,m x 500\,m zone. \mm{controllare che sia corretto} \lv{e' piu' complicato di cosi'...} 
Not reported here for the sake of brevity, we compare the  number of trips generated from the model and the ones presented in the original trace. As expected, there is a very good match with low residuals. We refer the reader to~\cite{ciociola2020} for details.

Notice that we employ a single global scaling factor $\lambda$ directly the temporal model to keep the spatial distribution of trips unchanged while increasing the request rate.

>>>>

%****************************************
% Motivation and outline
%****************************************
\subsection{Dataset}
In this chapter I leverage the data collected chapter \ref{chap:2_dataset} and described in chapters \ref{chap:3_charact} and \ref{chap:4_cs_comparison}, capturing real trips performed by car2go users. These data let us model the users' mobility demand in time and in space. From this, I derive a demand model that generalize the users' demand observed in the real data. I use it to generate realistic traces describing possible user trips and feed them to the event-based simulator to derive performance figures.

Recalling that the dataset consists in actual rental performed by car2go in Turin composed by geo-temporal coordinates recording where and when a user's ride starts and end. Figure~\ref{fig:10_3_rentals_per_day} summarizes the number of daily rentals from June 2017 to January 2018 in Turin, the reference dataset\footnote{For some periods the crawler did not record data due to server failures.} and table~\ref{tab:10_3_rental_data} outlines the main characteristics of this data. 400 cars were available, travelling on average less than 4\,km in each trip, for an average rental time of 21 minutes.

@@

%\subsection{Dataset}
%
%Our data consist in actual rentals performed by car2go users in Turin. Each observed rental has precise geo-spatial coordinates for trip origin and destination, and accurate timestamps. In Figure~\ref{fig:10_3_rentals_per_day} we summarize the number of daily rentals from June 2017 to January 2018 in Turin, our reference dataset\footnote{For some periods we did not record data due to server failures.}.  Despite the fact that rentals are non-stationary, especially during periods like August and Christmas holidays, the service usage follows a hourly and weekly pattern, not showing any particular growth. From this data, we select three months - from October 1st to December 31st.
%Table~\ref{tab:10_3_rental_data} outlines the main characteristics of this data. 400 cars were available, traveling on average less than 4\,km in each trip, for an average rental time of 21 minutes.
%
%
%To catch customers' habits, in Figure~\ref{fig:10_3_rentals_per_hour} we detail the average number of rentals per hour, separately per working-days (Monday to Friday) and per weekends. The weekdays hourly profile reflects the commuting pattern, with two clear peaks in the morning and evening rush hours. Conversely, during the weekends the number of night rentals is higher, likely due to more nightlife, while the morning peak is drastically smoothed.
%Not shown here for lack of space, our data allow us to observe the spatial diversity, with different origin and destination areas over the city. This shows how fundamental is to use actual data to build realistic scenarios for accurate system analysis.

\begin{table}[t]
\begin{center}
\setlength\tabcolsep{5pt} % default value: 6pt reduce cell padding
\caption{Main dataset characteristics, recorded in Turin from October to December 2017. Rental time and rental distance report both median (Med) and average (Avg) values.}
\begin{tabular}{|l|l|l|c|c|c|l|}
\hline
\!\multirow{3}{*}{Rentals}\!& \multirow{2}{*}{Fleet} &  \multicolumn{2}{|c|}{Rental}  &  \multicolumn{2}{|c|}{Rental}  &  \!\multirow{3}{*}{Zones}\!\\ 
& \multirow{2}{*}{Size}  & \multicolumn{2}{|c|}{Time~[min]}  & \multicolumn{2}{|c|}{Dist.~[km]} &    \\ \cline{3-6} %&  \multirow{2}{*}{Size[GB]}
 &  &\!Avg\!&\!Med\!&\!Avg\!&\!Med\!& \\ \hline
\hline
 180k & 400 & 21 & 20 & 3.96 & 3.36 & 279 \\ \hline
\end{tabular}
\label{tab:10_3_rental_data}
\end{center}
\end{table}

@@

%****************************************
% Demand model
%****************************************
%\subsection{Demand Model}

Instead, directly use the original trace to observe system performance, I decided to create a model to observe what-if scenarios, e.g., to observe the impact of a growth in the demand. For this, I use the available data to create a generalized demand model. I follow the approach presented in in~\cite{ciociola2020}. In a nutshell the demand is modelled in time by using modulated Poisson processes - a common accepted model for independent service requests of a very large population~\cite{poisson_inh}. To capture the spatial heterogeneity, I generalize the traces using Kernel Density Estimation (KDE)~\cite{kde_spatial}. KDE gives us the possibility to smooth the real data over a multi-dimensional space while maintaining the origin/destination correlation.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%Model of departure times
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
In more details, for the request arrival time process the model assumes that the inter-arrival time of trips follows an exponential distribution with rate depending on the type (weekend or working day) and hour of the day. The model considers 24 time bins of 1\,h each - 48 periods in total. In each time/day bin, the Poisson arrival rate matches the average rate of requests in that time bin in the original dataset as shown in figure~\ref{fig:10_3_rentals_per_hour}. This temporal model allows to scale the overall demand by introducing a global scaling factor $\lambda$ as a multiplier of the request rate of each time bin.

To model the spatial diversity of the demand, the models works on the case of study city divided in a set $Z$ of contiguous 500\,m x 500\,m zones, obtaining in total 279 zones. Each couple of spatial coordinates in the city area  ($x,y$) maps to one and only one zone.  Since each trip $i$ departs from a certain zone (origin $O_i$, described by two coordinates) and arrives to another zone (destination $D_j$, described by two coordinates), it is therefore characterized by two couples of coordinates, that can be represented as 4 scalars. 
For each time bin, I derive an origin and destination matrix counting how many trips were originated from a given zone $O$ and destined to a given zone $D$. Thus, in order to model the OD matrix in each temporal slot, I fit a 4-dimensional KDE based on the aforesaid coordinates.
For each of these matrices, I compute a KDE model, using Gaussian kernels, with bandwidth equal to 1. 
Not reported here for the sake of brevity, I compare the  number of trips generated from the model and the ones presented in the original trace. As expected, there is a very good match with low residuals. I refer the reader to~\cite{ciociola2020} for details.

Notice that I employ a single global scaling factor $\lambda$ directly the temporal model to keep the spatial distribution of trips unchanged while increasing the request rate.

### 2020-11-16 Chapter10/10_4-modeling.tex
chap_9_parsed

%In this section we briefly describe how we compute our demand model and the the assumption of our EFFCS simulator.
%In this work we leverage as input the data collected in our previous work~\cite{umap} describing real trips performed by car2go users. By using these data we model the users' demand in time by using modulated Poisson processes, and in space using Kernel by Density Estimation (KDE). 
%The aim of the demand model is to generalize the users' demand described in the real data. We use this generalized demand to generate samples describing user trips and feed them into our event-based simulator.

%\subsection{Dataset and Demand Model}

%Our input data consists in real trips performed by car2go users in 23 cities during 2017. While this input data are fundamental to understand the users' demand, we cannot directly use them to understand how the system scale with an increasing demand. As such, we use them to create a generalized demand model. For this, we follow the approach we presented in~\cite{its}. In a nutshell we model the demand in time by using modulated Poisson processes - a common accepted model for i.i.d. service requests of a very large population~\cite{poisson_inh}. While for space, we generalize the demand using Kernel Density Estimation (KDE)~\cite{kde_spatial}. KDE gives us the possibility to smooth the point process of the real data over a multi-dimensional space while maintaining the origin/destination correlation.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%Model of departure times
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\subsubsection{Time Modeling}

%For the time we assume that the inter-arrival time of trips follows an exponential distribution with  rate depending on the type and hour of the day. To account for the highly periodical rate as seen in Fig.\ref{fig:trips_day}, here we distinguish between weekday and weekends. We consider 24 time bins of 1\,h each (48 periods in total), where
%the Poisson arrival rate reflects the average rate of requests in the dataset. This allows to  scale the overall demand by introducing a global scaling factor.
%Not reported here for the sake of brevity, we compare the  number of trips in the simulated and the disaggregated trace. As expected, there is a very good match with low residuals (relative percentage errors for total trips between 0.6\% and 1.3\% for Louisville, 0.8\% and 3.4\% for Minneapolis)

Armed with the generalized demand model, we design and implement an event driven simulator to study the EVs FFCS system. Here we detail the simulation model, the simulator assumptions, the performance metrics, and the cost model used to project system performance into economic figures.

@@

We consider a fleet $F$ of electric cars. As before, we divide the city into a set $Z$ of zones of 500\,m x 500\,m each, where  cars can be parked, rented, charged and returned. Car characteristics are the same as MY2018 electric Smart ForTwo, namely $B=17.6$\,kWh battery capacity and $15.9$\,kWh/$100$\,km energy efficiency. Each car is characterized by its location, status (i.e., available, rented, under charge) and battery State of Charge (SoC). 
At simulation startup, cars are randomly placed in zones, with initial SoC uniformly distributed in $[0.5 B , B]$, and marked as available.

The charging infrastructure considers $n_p$ Level-2 chargers, with 3.7\,kW nominal power and 92\% charging efficiency. Charging stations are spread around the city zones. We place poles in those zones having the highest probability of being destination zones. This results in a good strategy to maximize system performance~\cite{7_cocca2019free,8_cocca2019free}. In details, we sort zones $z\in Z$ by the total number of parkings $tot\_park(z)$ observed in the original trace. We then consider the top $z_{p}$ fraction of zones, and place a number of poles in each proportionally to
$
%\left \lfloor
{n_p \cdot tot\_park(z)/\sum_z tot\_park(z)}
%\right \rfloor
$.

@@

\textbf{Car request event.}
When a \emph{rental requests} fires, a customer looks for a car within the origin zone and in 1-hop neighbouring zones.
If at least one car with enough SoC to reach the desired destination exists, the car gets rented, and a \emph{car release} event is scheduled after the time to reach the destination that is proportional to the distance, considering both orography and road network shape~\cite{8_cocca2019free}. If more than one such cars exists we choose the closest one, and, if need be, we choose the one with highest SoC.  
If no car is suitable for this ride, the trip does not occur and the request is marked as \emph{unsatisfied}.

@@

If instead the SoC is below $\alpha$, the car battery needs to be charged. The system handles the charging event by moving the car to the nearest-free charging pole. The simulator schedules a \emph{charge complete} event which accounts for both the time to reach the pole and the time to bring the SoC to 100\%\footnote{For simplicity, we assume there are infinite workers to handle the battery charge events so that a car gets serviced immediately. In case all poles are busy, the car gets placed in a queue of the closest charging station, and gets serviced when the first pole is freed.}.

\textbf{Charge complete.} When a \emph{charge complete} event fires, the car is marked as available, and customers can rent it again. The charging pole is released as well.  
Notice that we assume the car is released in the same zone where it was being charged, i.e., the system does not implement any relocation policy after charging.

@@

%****************************************
% Performance metrics
%****************************************
\subsection{Performance metrics}
In this work, we focus on the following performance metrics to compare different design options:

@@

%****************************************
% Cost model
%****************************************
\subsection{Cost model}
While performance indexes are useful to explore design options, the FFCS operator is ultimately interested in the economic sustainability of a solution. For this, we derive a cost model based on yearly projections.
We then consider revenues by projecting the number of rental and their duration. Armed with both, we estimate profit.
Here we consider:

\textbf{Vehicle cost.} We assume cars are leased to include all costs, namely registration, tax, insurance, ordinary and extraordinary maintenance, and roadside assistance. We assume electric cars do not pay for parking on street and for accessing limited traffic areas. Given the yearly car lease $C_{lease}$ and the number of vehicles, we easily derive the total yearly fleet cost.

\textbf{Charging infrastructure cost.} Here we refer to actual use cases as defined in~\cite{evInfrastructureCosts}. Pole installation costs account for material and labor cost. Material cost $C_{pole}$ includes hardware cost for Level II chargers.
Labor cost $C_{labor}$ is highly dependent on the city, region and country.
We need also to consider the make-ready infrastructure cost $C_{setup}$ that represents the cost for a charging station setup. It does not depend on the number of charging poles per station, but depends only on the number of charging zones $z_p\cdot|Z|$. It represents a highly variable cost since it depends on the location and the electric distribution infrastructure already in place. In fact, the expenses of trenching and laying conduit can add thousands of Euros to costs. %They are typically a large percentage of the capital costs of an installation, at about 30-40\%. 
All these costs are one-time costs. We assume these costs have an amortization period equal to the average charging station and pole lifetime $p_{life}$.

Next, we consider pole maintenance costs $C_{maint}$, which we derive from variable site-specific parameters.

In some cities, we need also to consider the per vehicle ground occupation tax $C_{ground}$, that usually depends on the surface for dedicated charging spot.
Due to the small size of Smart ForTwo, charging spots are assumed to be 4,50\,m x 2,30\,m, for each pole.

\textbf{Operating costs.} For this we take into account the $C_{energy}$ cost for the energy to charge the vehicles; the hourly cost for workers $C_{drivers}$ who have to handle the charge events; a cost $C_{disinf}$ to clean and disinfect the car any time the worker brings it to charge. Finally, we assume exterior car washing every 100 rentals, each costing $C_{wash}$.

\textbf{Rental Revenue} We consider a simple average cost-per-minute $R_{rental}$. This allows us to transform the total rental minutes into the total revenues.

Top part of Table~\ref{tab:10_4_summary} summarizes the parameters that define the scenarios used in the simulations. Bottom part shows the cost we consider for the Turin use case.
Given a scenario, we run a simulation to collect performance indexes. We next post-process the simulation results to derive the monthly cost and revenue figures. The custom simulator used is written in Python and based on SimPy library.\footnote{SimPy is a discrete-event simulation library. Documentation is available at: https://simpy.readthedocs.io/en/latest/contents.html.} The cost-revenue model is implemented in Python too and it is available online. The cost model allows one to interactively observe what happens by changing the cost values.\footnote{The code and data for cost and profit evaluation are available at: \url{https://smartdata.polito.it/on-scalability-of-electric-car-sharing-in-smart-cities/}.}

>>>>

Armed with the generalized demand model, I design and implement an event driven simulator, improving the software depicted in chapter \ref{chap:5_simulator} to study the EVs FFCS system. Here I detail the simulation model, the simulator assumptions, the performance metrics, and the cost model used to project system performance into economic figures.

@@

I consider a fleet $F$ of electric cars. As the old version, the simulator works on the city divided $Z$ of zones of 500\,m x 500\,m each, where cars can be parked, rented, charged and returned. Car characteristics are the same as MY2018 electric Smart ForTwo, namely $B=17.6$\,kWh battery capacity and $15.9$\,kWh/$100$\,km energy efficiency like the previous works. Each car is characterized by its location, status (i.e., available, rented, under charge) and battery State of Charge (SoC). 
At simulation startup, cars are randomly placed in zones and as first new feature, the initial SoC uniformly distributed in $[0.5 B , B]$, and marked as available.

The charging infrastructure considers $n_p$ Level-2 chargers, with 3.7\,kW nominal power and 92\% charging efficiency. Charging stations are spread around the city zones. The simulator places poles in those zones having the highest probability of being destination zones. This results in a good strategy to maximize system performance~\cite{7_cocca2019free,8_cocca2019free}. In details, the algorithm sorts zones $z\in Z$ by the total number of parkings $tot\_park(z)$ observed in the original trace. The simulator then consider the top $z_{p}$ fraction of zones, and place a number of poles in each proportionally to
$
%\left \lfloor
{n_p \cdot tot\_park(z)/\sum_z tot\_park(z)}
%\right \rfloor
$.

@@

\textbf{Car request event.}
When a \emph{rental requests} fires, a customer looks for a car within the origin zone and in 1-hop neighbouring zones.
If at least one car with enough SoC to reach the desired destination exists, the car gets rented, and a \emph{car release} event is scheduled after the time to reach the destination that is proportional to the distance, considering both orography and road network shape~\cite{8_cocca2019free}. If more than one such cars exists the closest one is picked, and, if need, the one with highest SoC.  
If no car is suitable for this ride, the trip does not occur and the request is marked as \emph{unsatisfied}.

@@

If instead the SoC is below $\alpha$, the car battery needs to be charged. The system handles the charging event by moving the car to the nearest-free charging pole. The simulator schedules a \emph{charge complete} event which accounts for both the time to reach the pole and the time to bring the SoC to 100\%\footnote{For simplicity, the assumption is the presence of infinite workers to handle the battery charge events so that a car gets serviced immediately. In case all poles are busy, the car gets placed in a queue of the closest charging station, and gets serviced when the first pole is freed.}.

\textbf{Charge complete.} When a \emph{charge complete} event fires, the car is marked as available, and customers can rent it again. The charging pole is released as well.  
Notice that I assume the car is released in the same zone where it was being charged, i.e., the system does not implement any relocation policy after charging.

@@

%****************************************
% Performance metrics
%****************************************
\subsection{Performance metrics}
In this chapter, I describe on the following performance metrics to compare different design options:

@@

%****************************************
% Cost model
%****************************************
\subsection{Cost model}
While performance indexes are useful to explore design options, the FFCS operator is ultimately interested in the economic sustainability of a solution. For this, I derive a cost model based on yearly projections.
I then consider revenues by projecting the number of rental and their duration. Armed with both, I estimate profit.
Here I consider:

\textbf{Vehicle cost.} I assume cars are leased to include all costs, namely registration, tax, insurance, ordinary and extraordinary maintenance, and roadside assistance. I assume electric cars do not pay for parking on street and for accessing limited traffic areas. Given the yearly car lease $C_{lease}$ and the number of vehicles, I easily derive the total yearly fleet cost.

\textbf{Charging infrastructure cost.} Here I refer to actual use cases as defined in~\cite{evInfrastructureCosts}. Pole installation costs account for material and labor cost. Material cost $C_{pole}$ includes hardware cost for Level II chargers.
Labor cost $C_{labor}$ is highly dependent on the city, region and country.
I need also to consider the make-ready infrastructure cost $C_{setup}$ that represents the cost for a charging station setup. It does not depend on the number of charging poles per station, but depends only on the number of charging zones $z_p\cdot|Z|$. It represents a highly variable cost since it depends on the location and the electric distribution infrastructure already in place. In fact, the expenses of trenching and laying conduit can add thousands of Euros to costs. %They are typically a large percentage of the capital costs of an installation, at about 30-40\%. 
All these costs are one-time costs. I assume these costs have an amortization period equal to the average charging station and pole lifetime $p_{life}$.

Next, I consider pole maintenance costs $C_{maint}$, which I derive from variable site-specific parameters.

In some cities, I need also to consider the per vehicle ground occupation tax $C_{ground}$, that usually depends on the surface for dedicated charging spot.
Due to the small size of Smart ForTwo, charging spots are assumed to be 4,50\,m x 2,30\,m, for each pole.

\textbf{Operating costs.} For this I take into account the $C_{energy}$ cost for the energy to charge the vehicles; the hourly cost for workers $C_{drivers}$ who have to handle the charge events; a cost $C_{disinf}$ to clean and disinfect the car any time the worker brings it to charge. Finally, I assume exterior car washing every 100 rentals, each costing $C_{wash}$.

\textbf{Rental Revenue} I consider a simple average cost-per-minute $R_{rental}$. This allows us to transform the total rental minutes into the total revenues.

Top part of table~\ref{tab:10_4_summary} summarizes the parameters that define the scenarios used in the simulations. Bottom part shows the cost I consider for the Turin use case.
Given a scenario, I run a simulation to collect performance indexes. I next post-process the simulation results to derive the monthly cost and revenue figures. The custom simulator used is written in Python and based on SimPy library.\footnote{SimPy is a discrete-event simulation library. Documentation is available at: https://simpy.readthedocs.io/en/latest/contents.html.} The cost-revenue model is implemented in Python too and it is available online. The cost model allows one to interactively observe what happens by changing the cost values.\footnote{The code and data for cost and profit evaluation are available at: \url{https://smartdata.polito.it/on-scalability-of-electric-car-sharing-in-smart-cities/}.}
